Practical Optional Types for Clojure

### 2015-02-23 formal-model.tex
isa example

We can also finally define \nil{} test

\begin{minted}{clojure}
(ann nil? (Pred nil))
(defn nil? [x]
  (isa? (class (class x)) Class))
\end{minted}

with the proposition set 
\isacompare{\s{}}{\classpe{}({\path{\classpe{}}{\x{}}})}{\Value{\Class}}{\filterset{\isprop{\Class}{\path{\classpe{}}{\x{}}}}{\notprop{\Class}{\path{\classpe{}}{\x{}}}}},

@@

For example, {\isaapp{\Keyword}{\Object}} = \true{} but 
{\isaapp{\Object}{\Keyword}} = \false{}.
This is useful in combination with \classconst{}
with a fixed second argument, where {\isaapp{\appexp{\classconst}{\x{}}}{\Keyword}}
is a test for \x{} being a \Keyword{}, and also
as an equality test like
{\isaapp{\x{}}{\makekw{a}}}.

We encode this behaviour in figure~\ref{main:figure:isatyping}. 
\isacompareliteral{} is the interesting part,
which takes the object of the first argument 
and the type of the second argument, and returns the then and else propositions
of the expression.

The first case handles subclassing mode
in expressions like {\isaapp{\appexp{\classconst}{\x{}}}{\Object}},
where the first argument has an object ending in \classpe{} and the second argument has a singleton class type.
For example, in this case \isacomparetwo{\path{\classpe{}}{\x{}}}{\Value{\Object}}
gives a then proposition {\isprop{\Object} {\x{}}}
and an else proposition {\notprop{\Object}{\x{}}}.

The reason to peel off one path element is subtle. For example,
claiming {\isprop{\Value{\Object}} {\path{\classpe{}}{\x{}}}}
might seem suitable, however it implies \inopenv{\openv{}}{\path{\classpe{}}{\x{}}}{\Object}
which is wrong if \x{} = \makekw{a} where \inopenv{\openv{}}{\path{\classpe{}}{\x{}}}{\Keyword}.
(Formulating \isacompareliteral revealed a three year old bug in the implementation).

The second case handles the equality mode of \isaliteral{}
with a non-class singleton type as the second argument, to be certain we are not in subclassing mode.
Substitution is used to ensure well formed propositions.

@@

\isaliteral{} serves as the default dispatch mechanism for multimethods in Clojure.

>>>>

We can use \isaliteral{} as a crude \nil{} test

\inputminted[firstline=5]{clojure}{code/demo/src/demo/isa_nil.clj}

because the then proposition is {\isprop{\Class}{\path{\classpe{}}{\x{}}}}, so
we know {\isprop{\Object}{\x{}}} by L-Update, and combined with the fact {\isprop{\Union{\Nil{}}{\Number}}{\x{}}}
we can infer {\isprop{\Number}{\x{}}} by L-Update.

@@

The second \isacompareliteral{} line handles the equality mode of \isaliteral{}
with a non-class singleton type as the second argument.

Substitution is used to ensure well formed propositions.

@@
